---
title: "task4"
author: "Anderson Monken"
date: "10/23/2020"
output:
  html_document: default
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
library(pacman)
p_load(tidyverse)
p_load(ggthemes)
p_load(lubridate)

# read in data
accounts_data <- read_csv('data/accounts_analytical.csv')
transactions_data <- read_csv('data/transactions.csv')

# find the account number with the most transactions
max_acct_transactions <- transactions_data %>% group_by(account_id) %>% count() %>% arrange(desc(n)) %>% head(1) %>% pull(account_id)

# get the transaction data on that account
df_max_acct <- transactions_data %>% filter(account_id == max_acct_transactions)
```

The first time series plot shows the balance of account number `r max_acct_transactions`

```{r}
# set up aes layer
ggplot(df_max_acct,
       aes(x = date,
           y = balance)) + 
  # geom layer
  geom_line() +
  # modify x axis
  scale_x_date(date_breaks = "1 year", date_labels ="%Y") + 
  # labels
  labs(title = paste("Time series of account balance\nfor account number",max_acct_transactions),
       x = NULL,
       y = "Account Balance") +
  # theme info
  theme_tufte() +
    theme(title = element_text(size = 16),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust = 0.5))


```


This plot is pretty noisy because there are a lot of transactions in short periods of time. If we aggregate to a monthly frequency, we can see how the average balance for each month results in a smoother curve. We notice the general decline in balance over time much more clearly at a monthly frequency than when we look at every balance.

```{r}
df_max_acct_monthly <- df_max_acct %>%
    group_by(year(date),month(date)) %>% 
  summarize(balance = mean(balance), date = max(date))

# set up aes layer
ggplot(df_max_acct_monthly,
       aes(x = date,
           y = balance)) + 
  # geom layer
  geom_line() +
  # modify x axis
  scale_x_date(date_breaks = "1 year", date_labels ="%Y") + 
  # labels
  labs(title = paste("Monthly time series of account balance\nfor account number",max_acct_transactions),
       x = NULL,
       y = "Account Balance") +
  # theme info
  theme_tufte() +
    theme(title = element_text(size = 16),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust = 0.5))

```

The next plot shows the credit and debit activity of the account to see if there were changes in the number of transactions made.

```{r, message = FALSE}
# calculate the number of debit and credit transactions for each month
df_account_usage <- df_max_acct %>% 
  group_by(year(date),month(date)) %>% 
  summarize(sum_credit = sum(type == "credit"), 
            sum_debit = sum(type == "debit"),
            date = mean(date)) %>%
  ungroup() %>%
  select(date, sum_debit, sum_credit) %>%
  pivot_longer(-date)


# set up aes layer
ggplot(df_account_usage,
       aes(x = date,
           y = value,
           color = name)) + 
  # geom layer
  geom_line() +
  # modify x axis
  scale_x_date(date_breaks = "1 year", date_labels ="%Y") + 
  # modify fill axis
  scale_color_discrete(breaks = c('sum_credit','sum_debit'), labels = c('Credit (deposit)','Debit (withdrawal)')) +
  # labels
  labs(title = paste("Monthly time series of account activties\nfor account number",max_acct_transactions),
       x = NULL,
       y = "Number of Transactions in a Month",
       color = 'Transaction Type') +
  # theme info
  theme_tufte() +
  theme(title = element_text(size = 16),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust = 0.5))

```


The final plot shows the total debit transaction amount for each method and category breakdown over time on a monthly frequency. The case transaction are the most variable for account 8261. If there is a second month with a high transaction volume in cash, the bank might consider flagging it since it only tends to happen every few months.

```{r}
# calculate the number of debit and credit transactions for each month
df_account_usage <- df_max_acct %>% 
  filter(type == 'debit') %>%
  mutate(category = if_else(is.na(category),'NA',category)) %>%
  group_by(year(date),month(date), category, method) %>% 
  summarize(tot_trans = sum(amount), 
            date = mean(date)) %>%
  ungroup() %>%
  select(date, method, category, tot_trans) %>%
  mutate(method_category = paste(method,'-',category))


# set up aes layer
ggplot(df_account_usage,
       aes(x = date,
           y = tot_trans,
           color = method_category)) + 
  # geom layer
  geom_line() +
  # modify x axis
  scale_x_date(date_breaks = "1 year", date_labels ="%Y") + 
  # labels
  labs(title = paste("Monthly time series of total transaction amount \nby type for account number",max_acct_transactions),
       x = NULL,
       y = "Total Transaction Amount (Monthly)",
       color = 'Transaction Type\n(method - category)') +
  # theme info
  theme_tufte() +
  theme(title = element_text(size = 16),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust = 0.5))
```

